---
title: "Copy_Vaccine"
author: "Sylvia Liu"
date: "16/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(tibble)
library(ggplot2)
library(ggthemes)
library(maps)

library(plotly) # Interactive data visualizations
library(viridis) # Color gradients
library(lubridate)

library(randomForest) 
library(ranger)       # a faster implementation of randomForest
library(caret)        # performe many machine learning models
library(broom)  # try: `install.packages("backports")` if diretly installing broom gives an error

library(reshape2) # to use melt()
```

## Reading data
```{r}
covid_data <- read.csv("owid-covid-data-160222.csv")
covid_data$date <- as.Date(covid_data$date, format = "%d/%m/%y")
#head(covid_data)
```

## Vaccinations progress  

### EDA - total_vaccinations
```{r}
# Selecting countries with enough vaccination data
# Set a baseline that the country we selected should has at least 180 days(i.e. 6 months) recorded for the value of vaccinations
# Set a baseline that the country we selected should has at least 120 days(i.e. 4 months) recorded for the value of vaccinations during 2021-4-1 to 2021-10-31


valid_countries_df <- covid_data %>%  
    filter(!is.na(total_vaccinations) & date<=as.Date("2021-10-31") & date >= "2021-04-01") %>%
    group_by(iso_code,location) %>% 
    dplyr::summarise(count=n())  %>% 
    filter(count >= 105 & !grepl("^OWID.", iso_code))  %>% # Remove continent and income groups
    ungroup()

valid_countries_df <- covid_data %>%  
     filter(!is.na(total_vaccinations) & !is.na(population)) %>%
     group_by(iso_code,location) %>% 
    dplyr::summarise(count=n())  %>% 
     filter(count >= 180 & !grepl("^OWID.", iso_code))  %>% ungroup()# Remove continent and income groups

    
selected_countries <- valid_countries_df$location
# selected_countries

filter_avaiable_date <- function(country) {
  data <- covid_data %>% filter(location == country) %>% filter(!is.na(total_vaccinations) & total_vaccinations != 0) %>% arrange(date)
  start_date <- data %>% select(date) %>% slice(1L) %>% pull()
  end_date <- data %>% select(date) %>% slice_tail() %>% pull()
  p <- ggplot(data, aes(x = date, y = total_vaccinations)) +
    geom_point()+ 
    geom_line(aes(x = date, y = total_vaccinations)) +
    ggtitle(country)+
    scale_x_date(limits = as.Date(c("2020-12-25","2022-2-22")))
  
  return(list(start_date=as.Date(start_date), end_date=as.Date(end_date), p=p))
}

results_d <- c(NA,NA,NA) #c("Country name", start_date, end_date)
for (i in selected_countries) {
  res <- filter_avaiable_date(i)
  start_date <- res$start_date
  end_date <- res$end_date
  results_d <- rbind(results_d,c(i,start_date,end_date))
  print(res$p)
}
```
> From the plot we can observed that，in generl, the overlapping time period for the vaccine surge in each country is April 2021 to October 2021, a total of eight months. We will use this time period as a criterion for subsequent modelling on VRI and factors.

### Vaccine Roll-Out Index (VRI) values for the countries

(Reference: https://www.mdpi.com/2076-393X/10/2/194/htm)

> **Estimate Vaccine Uptakes Rates r* for Countries**
"The logistic growth model was used to estimate the vaccination uptake rate. In the logistic model, the cumulative number of doses administered c(t) satisfies the following equation: $$c(t) = \frac{K}{1+a\cdot{e}^{-r\cdot t}}$$,"_ where K is the total vaccinations administered at the end of the date recorded, r is the vaccination uptake rate(i.e. speed), and $\frac{K}{1+a} = c(0)$ is the initial number of doses given.

```{r}

set.seed(3888)

### Estimate Vaccine Uptakes Rates r* for Countries
# Step 1: A function that returns the logistic model used for estimating r*

cutoff_date_start = "2021-4-01" # As described in the EDA section
cutoff_date_end = "2021-10-31" # As described in the EDA section

estimate_uptake_rate <- function(country) {
  
  data <- covid_data %>% filter(location == country) %>% filter(!is.na(total_vaccinations) & total_vaccinations != 0) %>% arrange(date) %>% filter(date <= as.Date(cutoff_date_end) & date >= cutoff_date_start) 
  K <- data %>% select(total_vaccinations) %>% slice_tail() %>% pull()
  
  ## Calculate t by ranking the date
  data$t <- rank(data$date)
      
  ## Taking the model to get fitted
  x <- data$t-1
  y <- data$total_vaccinations
  m <- nls( y ~ (K / (1 + a*exp(-r*x))), start = list(a=0.01,r=0.01),control=nls.control(maxiter=200)) #a,r are the coefficients    
  
  return(list(model=m))
  
}

# Example
res_nls  = estimate_uptake_rate("Japan")
summary(res_nls$m)
```

**Evaluation of the metric on est.r**
```{r}
## Step2: Another function for evaluating the logistic model we created for the estimate r (r*), 
## including fitted plots, correlation values and residual values
evaluate_est_r <- function(country, m) {
  
  data <- covid_data %>% 
    filter(location == country) %>% 
    filter(!is.na(total_vaccinations) & total_vaccinations != 0) %>% 
    arrange(date)  %>% filter(date <= as.Date(cutoff_date_end) & date >= cutoff_date_start) 
  
  # goodness of fit
  corr = round(cor(data$total_vaccinations, predict(m)),4)
     
  # minimized residual value
  resid = round(sum(resid(m)^2),4)
     
  # Plot fitted values vs true values
  p <- ggplot(data, aes(x = date, y = total_vaccinations)) +
  geom_point()+
  geom_line(aes(x = date, y = total_vaccinations), color="green")+
  geom_line(aes(x=date,y=predict(m)), color="blue") 
  
  return(list(plot=p,corr=corr,resid=resid))
}

# Example:
eval_res  = evaluate_est_r("Japan",res_nls$m)
eval_res$plot
paste("goodness of fit:",eval_res$corr)
paste("minimized residual value",eval_res$resid)

```

> Define and Calculate Vaccine Roll-Out Index (VRI) 
"The Vaccine Roll-out Index (VRI) for a country was defined as follows: $$VRI=r\cdot \frac{d}{N}$$," where r is the vaccination uptake rate as defined in the previous section, d is the total vaccinations, and N is the population. 

**VRI was used as an index to compare the overall vaccination progression among different countries, as it reflects both the speed and the density/extend.**

```{r}
# Step 3 calculate VRI using: est_rate * d/N
## TODO: adding constraints on spec_date, can only have date in range 2021-04-01 ~ 2021-10-31
cal_VRI <- function(country, est_rate, spec_date) {
  data <- covid_data %>%  filter(location == country) %>% filter(!is.na(total_vaccinations) & total_vaccinations != 0) %>% arrange(date) %>% filter(date <= as.Date(spec_date) & date >= cutoff_date_start) 
  if (nrow(data) != 0) {
    d <- data %>% select(total_vaccinations) %>% slice_tail() %>% pull()
    N <- data %>% filter(!is.na(population)) %>% select(population) %>% slice_tail() %>% pull()
    VRI = est_rate * d/N
  }
  else {  ## there's no total_vaccinations data recorded ahead of this time
    VRI = 0
  }
 
  return(vri=VRI)
  
}


```


> Comment: The function could be adjusted accordingly for the shiny app:
- Changing a single end date to a range of dates.


_Side Note: The date user entered should be in between the_ `start_date` _and_ `end_date`.
### Calculate `est.r` and `VRI` for all countries with no empty values and 0s in `toal_vaccinations` column within the input date
```{r}

# Calculate est_r, using the whole dataset available to estimate r* (more precise)
results_r <- c(NA,NA) #c("Country name", est_r)
for (i in selected_countries) {
  res_model <- estimate_uptake_rate(i)$model
  results_r <- rbind(results_r,c(i,round(summary(res_model)$parameters[2],4)))
}
results_r <- results_r[-1,]
colnames(results_r) <- c("Location","Est.r")
results_r <- data.frame(results_r)
results_r$Est.r <- as.double(results_r$Est.r)
#head(results_r)
results_r
```


### The association between vaccination uptake rate and social-economic factors

#### Merging datasets

**I have removed the government trust dataset as it only contains OECD countries**
```{r}

corruption <- read_xlsx("CPI2020_GlobalTablesTS_210125.xlsx", sheet=1,skip=2)
corruption <- corruption %>% select (c("ISO3", "CPI score 2020")) 
#corruption

happiness <- read_csv("happiness-cantril-ladder.csv")
happiness <- happiness %>% group_by(Code) %>%
  arrange(Year) %>%
  filter(!is.na(`Life satisfaction in Cantril Ladder (World Happiness Report 2021)`)) %>%
  dplyr::summarise(satisfaction=last(`Life satisfaction in Cantril Ladder (World Happiness Report 2021)`)) 

covid_grouped <- covid_data  %>% group_by(iso_code) %>% filter(location %in% selected_countries) %>% 
  arrange(date) %>% 
  slice_tail()

joint_data <- merge(covid_grouped,corruption,  by.x=c("iso_code"), by.y=c("ISO3"))

joint_data <- merge(joint_data,happiness, by.x=c("iso_code"), by.y=c("Code"))
joint_data

# GHE-INDEX
ghs <- read_csv("2021-GHS-Index-April-2022.csv")
ghs <- ghs %>% filter(Year == 2021) %>% select(Country,`OVERALL SCORE`)
merge(joint_data, ghs,  by.x=c("location"), by.y=c("Country"))

joint_data <- merge(joint_data, ghs,  by.x=c("location"), by.y=c("Country"))
colnames(joint_data)[70] <- "GHS_score"
joint_data
```

Merge joint covid data with VRI.
```{r}

results_vri <- c(NA,NA) #c("Country name", VRI)

for (i in c(1:nrow(results_r))) {
  vri = cal_VRI(results_r$Location[i],results_r$Est.r[i], cutoff_date_end)
  results_vri <- rbind(results_vri,c(results_r$Location[i],round(vri,4)))
}
results_vri <- results_vri[-1,] # removing the first pair of NAs
colnames(results_vri) <- c("Location","VRI")
results_vri <- data.frame(results_vri)
results_r$VRI <- as.double(results_vri$VRI)
results_vri

# joint_data
vri_covid_cov_data <- merge(joint_data,results_vri,by.x=c("location"), by.y=c("Location"))
vri_covid_cov_data$VRI <- as.double(vri_covid_cov_data$VRI)
vri_covid_cov_data
```


#### Selecting Covariates

##### Demographic factors (scatter plots)
Variables: `population_density`, age(`aged_65_older`/`aged_70_older`), `gdp_per_capita`, 

```{r}
# Basic scatter plot
df <- vri_covid_cov_data %>% filter((VRI < 0.07)& !is.na(population_density) # set a threshold to remove outlier
                                    &!is.na(gdp_per_capita) &!is.na(aged_65_older)&
                                      !is.na(extreme_poverty)&
                                      !is.na(human_development_index)& 
                                      !is.na(cardiovasc_death_rate)& 
                                      !is.na(diabetes_prevalence)&
                                      !is.na(hospital_beds_per_thousand)) %>% 
                                      
                            select(c(location,VRI,population_density,gdp_per_capita,aged_65_older,
                                     `CPI score 2020`,
                                     satisfaction,extreme_poverty,human_development_index,life_expectancy,
                                      cardiovasc_death_rate,diabetes_prevalence,
                                      hospital_beds_per_thousand, GHS_score)) 
df
```


```{r}
ggplot(df, aes(x=VRI, y=population_density)) + geom_point() + ggtitle("Population density vs Vaccine Roll-out Index(VRI)") + ylim(c(0,500))
ggplot(df, aes(x=VRI, y=gdp_per_capita)) + geom_point() + ggtitle("GDP per capita vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y=aged_65_older)) + geom_point() + ggtitle("Aged over 65 vs Vaccine Roll-out Index(VRI)")
```

##### Social and Economic
Variables: Stringency(`stringency_index`), trust(`share_who _trust_government`), corruption(`CPI score 2020`), happiness(`satisfaction`), `extreme_poverty`, `human_development_index`, `life_expectancy`.

Remove the stringency_index as there's not enough data.
```{r}
#ggplot(df, aes(x=VRI, y= stringency_index)) + geom_point() + ggtitle("Stringency vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y=`CPI score 2020`)) + geom_point() + ggtitle("Corruption vs Vaccine Roll-out Index(VRI)")
#ggplot(df, aes(x=VRI, y= `Trust the national government in this country`)) + geom_point() + ggtitle("Trust in government vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y= satisfaction)) + geom_point() + ggtitle("Happiness vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y= extreme_poverty)) + geom_point() + ggtitle("Extreme Poverty vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y= human_development_index)) + geom_point() + ggtitle("Human Development Index vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y= life_expectancy)) + geom_point() + ggtitle("Life Expectancy vs Vaccine Roll-out Index(VRI)")

```

##### Health
variables: `cardiovasc_death_rate`/`diabetes_prevalence`, smokers?
```{r}
ggplot(df, aes(x=VRI, y= cardiovasc_death_rate)) + geom_point() + ggtitle("Cardiovasc Death Rate vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y= diabetes_prevalence)) + geom_point() + ggtitle("Diabetes Prevalence vs Vaccine Roll-out Index(VRI)")
#hospital_beds_per_thousand
ggplot(df, aes(x=VRI, y= hospital_beds_per_thousand)) + geom_point() + ggtitle("hospital_beds_per_thousand vs Vaccine Roll-out Index(VRI)")
ggplot(df, aes(x=VRI, y= GHS_score)) + geom_point() + ggtitle("GHS vs Vaccine Roll-out Index(VRI)")
```

##### Correlation Check
```{r}
# correlation matrix
cor_mat <- round(cor(df %>% select(-location),method="spearman"),2)# for non-linear correlation (non-parametric)

## Reorder the correlation matrix
# Refernce: http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization

reorder_cormat <- function(cor_mat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cor_mat)/2)
  hc <- hclust(dd)
  return (cor_mat[hc$order, hc$order])
}

```


```{r}
## Plotting heatmap
plot_heatmap <- function(cor_mat){
  
  cor_mat <- reorder_cormat(cor_mat)

  ## Remove redundant information (taking the upper triangles values)
  cor_mat[lower.tri(cor_mat)]<- NA
  cor_mat <- melt(cor_mat, na.rm = TRUE)
  
  ggheatmap <- ggplot(cor_mat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Spearman\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 8, hjust = 1))+
 coord_fixed()

  ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 3) +
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))

}

plot_heatmap(cor_mat)

```

#### Modeeling (RF)

https://topepo.github.io/caret/recursive-feature-elimination.html#backwards-selection

##### Check correlation
```{r}
df <- df %>% select(-location)
cor_mat <- round(cor(df,method="spearman"),2)# for non-linear correlation
## Remove redundant information (taking the upper triangles values)
cor_mat[lower.tri(cor_mat)]<- NA
cor_mat <- melt(cor_mat, na.rm = TRUE)
as.data.frame(cor_mat) %>% filter(value > 0.8 & value != 1)
```



##### Split data into training and testing
```{r}
ind <- createDataPartition(df$VRI, p = .80,list=FALSE) 
train <- df[ind,]
X_test <- df[-ind,] %>% select(-VRI)
y_test <- df[-ind,] %>% select(VRI)
dim(train)
dim(X_test)
train

```


##### Tune models on training data with all predictors using grid search with cross-validation
```{r warning=FALSE}
# REname for Random Forest 
colnames(df)[5] <- "CPI_score_2020" 

# hyper parameter grid search (definitely need a bit modify)
mtry <-  seq(3, 11, by = 1)
num_trees <- c(100,150,200,250,300,350,400,450,500)


# Manual Search
#control <- trainControl(method="cv", number=3, search="grid")
grid_param <- expand.grid(.mtry=mtry)
modellist <- list()
for (ntree in num_trees) {
	set.seed(3888)
	fit <- train( VRI~., 
                      data= train %>% select(-human_development_index) , 
                      method='rf', 
                      tuneGrid=grid_param, 
	                    ntree= ntree,
                      trControl=trainControl(method='cv', 
                        number=3) )
	key <- toString(ntree)
	modellist[[key]] <- fit

}



# compare results 
lowest_mse <- 1
model_mse <- modellist[[1]]
highest_r2 <- 0
model_r2 <- modellist[[1]]
for (i in c(1:length(modellist))) {
  #print(modellist[[i]]$finalModel)

  mse = mean(modellist[[i]]$finalModel$mse)
  r2 = mean(modellist[[i]]$finalModel$rsq)
  if (lowest_mse > mse){
    lowest_mse = mse
    model_mse = modellist[[i]]$finalModel
    }
  if (highest_r2 < r2){
     highest_r2 = r2
     model_r2 = modellist[[i]]$finalModel
   }
}
model_mse
model_r2 


```

###### Predict on test sample
```{r}
predictions <- predict(model_mse,newdata = model.matrix(~.,X_test))
# MSE
sum((y_test-predictions)^2) / length(y_test)
# R squared
as.double(cor(y_test, as.data.frame(predictions)) ^ 2)
```


###### Backward Selection(RFE - recursive-feature-elimination)
```{r}
set.seed(3888)

subsets <- c(3:11)

ctrl <- rfeControl(functions = rfFuncs, # using random forest
                   method = "cv",
                   number = 3, # using 3-fold cross validation
                   verbose = FALSE)

rfProfile <- rfe((train %>% select(-VRI,-human_development_index)), as.matrix(train %>% select(VRI)),
                 sizes = subsets,
                 rfeControl = ctrl)

rfProfile
as.data.frame(predictors(rfProfile))
rfProfile$fit
trellis.par.set(caretTheme())
plot(rfProfile, type = c("g", "o"))

```

Optimal Model for RF predict on test data
```{r}
rfProfile
###### Predict on test sample
predictions <- predict(rfProfile,newdata = X_test)
# MSE
sum((y_test-predictions)^2) / length(y_test)
# R squared
as.double(cor(y_test, as.data.frame(predictions)) ^ 2)
```

```{r}
rf_df <- df %>% select(VRI,life_expectancy, extreme_poverty, cardiovasc_death_rate, gdp_per_capita)

rf <- randomForest(VRI ~ ., 
                data= rf_df, ntree=500,mtry=2,
                keep.forest=FALSE, importance=TRUE, seed=3888)

vimp <- varImp(rf,scale = TRUE)
vimp <- vimp[order(vimp$Overall,decreasing = TRUE),,drop = FALSE]
vimp$var <- rownames(vimp)                  
vimp$Overall <- round(vimp$Overall,3)

options(scipen=999)
vimp %>%
  arrange(desc(Overall)) %>%
  ggplot(aes(reorder(var, Overall), Overall)) +
  geom_col() +
  coord_flip() +
  ggtitle("Top important variables") +
  xlab("Factors in order") +
  ylab("Scaled importance")

```

##### Modelling Linear Regression (backwards selection using single p-value
i.e. using drop1() from the full model.
```{r}
# Check strong correlated predictors
cor_mat <- round(cor(df %>% select(-VRI)),2) # for linear correlation
cor_mat[lower.tri(cor_mat)]<- NA
cor_mat <- melt(cor_mat, na.rm = TRUE)
as.data.frame(cor_mat) %>% filter(value > 0.8 & value != 1)

# Full model (Removing non-linear trend predictors + strong correlated predictors)
M1 = lm(VRI ~ ., data = train %>% select(-extreme_poverty,-population_density))  # ,-human_development_index,-`CPI score 2020`

# Try doing backward selection
step_back_aic = step(M1, direction = "backward", trace = FALSE)
summary(step_back_aic)


```


Check assumptions
```{r}
library(ggfortify)
autoplot(step_back_aic, which = 1:2) + theme_bw()
```
Predict
```{r}
predictions <- predict(step_back_aic, newdata = X_test)
# MSE
sum((y_test-predictions)^2) / length(y_test)
# R squared
as.double(cor(y_test, as.data.frame(predictions)) ^ 2)
```