---
title: "DATA3888 COVID Vaccination"
subtitle: "Group C7"
author: "Aurora Wang, David Vucic, Diana Cao, Harun Hodzic, Sylvia Liu, Tina Tian, Vighnesh Deshpande"
institute: "University of Sydney"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.align = "center"
)
```


```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#43418A", secondary_color = "#03A696",
  header_font_google = google_font('Libre Baskerville', "400"),
  text_font_google   = google_font("DM Sans", "400", "400i"),
  code_font_google   = google_font("IBM Plex Mono")
)
```



# Introduction (30s)

<!-- Feel free to change the content. I am not entirely sure about this part -Aurora -->

.pull-left[
### Background
  - Data set
    - OWID COVID Data set
    - COVID Availability Policy
    - 2021 GHS Index
    - Corruption Perceptions Index 2020
    - World Life satisfaction 2021
]
  
  
.pull-right[

### Motivation

Develop an index to measure the vaccination uptake speed and coverage among countries. Then find out the factors that may highly affect this index.

]

.center[
### Investigate factors that may affect the vaccination uptake

  <!-- Vaccine Roll-Out Index -->
  
  <!-- Vaccination Time lag -->
  
]
---

# Vaccination Time Lag: First dose vs Fully vaccination (40s)

- highlight & Innovation
- Evaluation strategies
  - include results


---

# Vaccination Roll-Out Index (100s)

- highlight & Innovation
- Evaluation strategies
  - include results

---

# Consultation (10s)

- Main results
- most of the finding shown in shiny app

---

# Demo (60s)