---
title: "DATA3888 COVID Report"
author: '500254239'
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  bookdown::html_document2:
    theme: 
      bootswatch: lumen
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
bibliography: [references.bib, references-pkg.bib]
csl: apa-6th-edition.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.cap = ""
)
```

```{r libraries}
library(tidyverse)
library(printr)
```

# Guide for RMarkdown report (remove for final report)

## Bibliographies in RMarkdown

- `references.bib`: for BibTeX of all sources you wish to reference
- `references-pkg.bib`: for BibTeX of all packages used in the report

Bibliographies will be included automatically at the end before appendix. Copy paste `bib` or `BibTeX` text from your sources into the file `references.bib` following the example format, leaving an empty line between entries. Name it however you want in the first line after `@document-format{` like in the example. For any package you are using, if there isn't already an entry for it in the bib file, then use the function `citation("package-name")` in R console and copy the `BibTeX` entry from the output, paste it into `references-pkg.bib`.

Examples of how to reference bib entries in-text (check the rmd file):

- Garth reference: [@RN83]
- `tidyverse` reference: [@tidyverse]

## Figures cross-referencing

Here is an example plot with caption and how to reference it in-text:

Use code chunk names to reference the figure output of that code chunk: Figure \@ref(fig:example-plot). *Note this will only show the number (default format: section-num.figure-num), you have to put "Figure" before the \@ref(fig:..) as well.*

```{r example-plot, fig.cap="Example plot"}
# example plot
mtcars %>% 
  rownames_to_column("car") %>% 
  data_frame() %>% 
  ggplot() +
  aes(x = factor(cyl), y = disp, colour = factor(cyl)) +
  geom_boxplot()
```


# Executive Summary



# Data Background


# Methods


## Time lag between 1st and 2nd dose


## Vaccine rollout index (VRI) and variable importance


## Vaccine uptake and availability


# Results


# Discussion & Conclusion


# References

::: {#refs}
:::

# (APPENDIX) Appendix {-}